### 2018-04-09 ident/docs/draft_2.tex
changes to sec 2 of results and changed v5 ident figure

	\section{Results:}
	First, in section \ref{sec:example}, we demonstrate the use of the methodology that we described in section \ref{sec:ident} to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in section \ref{sec:small-model}. Identifiability results for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ are also discussed in this section.

	First in section \ref{sec:closedform}, we show that the ability to establish practical identifiability of kinetic parameters in metabolic network models using our methodology, described in section \ref{sec:ident}, relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes. We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.

@@

	Then, in section \ref{sec:data_utility} we provide a motivation for the need for experimental design, especially to identify kinetic models of metabolic networks. We do this by looking at the number of combinations that can identify the maximum number of parameters. 
	In section \ref{sec:trends} we show that the degree of identifiability of the maximum reaction rate parameter is always higher than the degree of identifiability of the corresponding enzyme binding parameter irrespective of enzyme kinetic rate law used to describe the flux. 
	In section \ref{sec:experiment_type} we discuss the ability to determine the type of experiments for parameter identification for a given flux based on the informativeness of a given type of experiment. In this section we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identification of the parameters for a given flux. Finally, we discuss some results arising out of the use of our methodology to determine the identifiability of parameters when data with additive noise is used in section \ref{sec:noise}.

@@

@@

	Similarly, when we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system. However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$. Thus, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have describe in section \ref{sec:ident}. Even in this scenario, just as we found for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable. $V_e^{max}$ on the other hand has only one possible root, thereby it is a globally identifiable parameter. The identifiability of the second root for $K_e^{fdp}$, indicated as $K_e^{fdp}(2)$, is shown in Figure \ref{fig:ident}d. The identifiability of the first root of $K_e^{fdp}$, $K_e^{fdp}(1)$, is shown in Figure \ref{fig:ident_root2}c.

	While the identifiability of the second roots $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$ for $v_3$ (Figure \ref{fig:ident_root2}b) are similar to that of the first root (Figure \ref{fig:ident}c), this is not the case for $v_5$. We would like to point out that this is one of the scenarios where the local identifiability of a parameter like $K_e^{fdp}$ can be expanded to global identifiability by adding physical relevance of parameter values as an additional constraint for practical identifiability ($D_k>0$, see Figure \ref{fig:ident-flowchart}b and section \ref{sec:ident}). We found that the first root for $K_e^{fdp}$, $K_e^{fdp}(1)$, never possessed a physically relevant value, and accordingly was never identifiable based on the criteria set out in Figure \ref{fig:ident}b and section \ref{sec:ident} (Figure \ref{fig:ident_root2}c). In contrast, the second root, $K_e^{fdp}(2)$ has an identifiability of 50.0\% (Figure \ref{fig:ident}d). Thus, we can get globally identifiable parameters through the addition of constraints related to their physical relevance to locally identifiable parameters.

@@

	we demonstrate a specific case of lack of persistence of excitation and the subsequent experimental design required to overcome this hurdle for parameter identifiability for a small nonlinear metabolic network model. Earlier, we mentioned that the informativeness of experiments for identifying parameters in a given flux depends on its ability to satisfy the conditions determined for practical identifiability of that parameter. For example, we demonstrated in section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them (Equation \ref{eq:v1_par}). Thus, in this instance, the informativeness of the experiments (changes in the input acetate concentration and $v_1$) for identifying parameters of $v_1$ is determined by the ability of the input change to effect a change in the measured value of $v_1$. Similarly, the identification of parameters for $v_2$ requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}. %, given both these variables are outputs of the model describing $v_2$.

	As such, as per Figure \ref{fig:figure1}a only 20\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$. Whereas, more than 90\% of the available data combinations satisfy the requirements for identifying $v_2$ (Figure \ref{fig:figure1}b). Thus, identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for both $v_1$ and $v_2$. Recall that we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Section \ref{sec:experiments}).
	In order to determine the type of experiments that help satisfy identifiability conditions for both parameters of $v_1$, we look at the distribution of the five different types in all data combinations that can identify these 2 parameters (Figure \ref{fig:figure2}a and b).

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1/v1_v2_v3_v5_ident}}
		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$ \textcolor{red}{(need to change figure)} when there is no noise in the input experimental data. The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter. The total number of experiments used to identify parameters in $v_1$, $v_2$ and $v_5$ is two. The total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is 240, 421 different combinations are used to test identifiability for $v_5$, and 4896 combinations are used to test identifiability of parameters for $v_3$. Note that $v_3$ is modeled with three parameters and requires data from at least three different experiments, and also has two possible roots for each of its parameters. Results for only one of the two roots indicated by (1) is shown here.}\label{fig:ident}
	\end{figure}

>>>>

	\section{Results:}
	First, in section \ref{sec:example}, we demonstrate the use of the methodology that we described in section \ref{sec:ident} to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in section \ref{sec:small-model}. Identifiability results for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ are also discussed in this section. We show that the ability to establish practical identifiability of kinetic parameters in metabolic network models relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes. We also show that the degree of identifiability of the maximum reaction rate parameter is always higher than the degree of identifiability of the corresponding enzyme binding parameter irrespective of enzyme kinetic rate law used to describe the flux.

	First in section \ref{sec:closedform},  We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.

@@

	Then, in section \ref{sec:data_utility} we provide a motivation for the need for experimental design, especially to identify kinetic models of metabolic networks. We do this by looking at the number of combinations that can identify the maximum number of parameters. 
	In section \ref{sec:trends}

	In section \ref{sec:experiment_type} we discuss the ability to determine the type of experiments for parameter identification for a given flux based on the informativeness of a given type of experiment. In this section we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identification of the parameters for a given flux. Finally, we discuss some results arising out of the use of our methodology to determine the identifiability of parameters when data with additive noise is used in section \ref{sec:noise}.

@@

	In Figure \ref{fig:ident}a less than 50\% of all data combinations can identify $v_1$. Whereas, all available data combinations satisfy the requirements for identifying $v_2$ (Figure \ref{fig:ident}b). This difference can be alluded to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter. For example we showed in Equation (\ref{eq:v1_par}) that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them. So, in this instance, the ability of a combination of experiments to be useful is determined by the ability of the input change (changes in the input acetate concentration) to effect a change in the measured value of $v_1$. Similarly, the identification of parameters for $v_2$ (Figure \ref{fig:ident}b) requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}.

@@

	Similarly, when we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system. However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$. Thus, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}. Even in this scenario, just as we saw earlier for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable. $V_e^{max}$ on the other hand has only one possible root, thereby it is a globally identifiable parameter. The identifiability of the second root for $K_e^{fdp}$, indicated as $K_e^{fdp}(2)$, is shown in Figure \ref{fig:ident}d. The identifiability of the first root of $K_e^{fdp}$, $K_e^{fdp}(1)$, is shown in Figure \ref{fig:ident_root2}c.

	While the identifiability of the second roots $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$ for $v_3$ (Figure \ref{fig:ident_root2}b) are similar to that of the first root (Figure \ref{fig:ident}c), this is not the case for $v_5$. We would like to point out that this is one of the scenarios where the local identifiability of a parameter like $K_e^{fdp}$ can be expanded to global identifiability by adding physical relevance of parameter values as an additional constraint for practical identifiability ($D_k>0$, see Figure \ref{fig:ident-flowchart}b and section \ref{sec:ident}). We found that the first root for $K_e^{fdp}$, $K_e^{fdp}(1)$, never possessed a physically relevant value, and accordingly was never identifiable based on the criteria set out in Figure \ref{fig:ident-flowchart}b and section \ref{sec:ident} (Figure \ref{fig:ident_root2}c). In contrast, the second root, $K_e^{fdp}(2)$ has an identifiability of 50.0\% (Figure \ref{fig:ident}d). Thus, we can get globally identifiable parameters through the addition of constraints related to their physical relevance to locally identifiable parameters.

@@

	We demonstrated in section \ref{sec:example}, using $v_1$ and $v_2$ as examples, that the ability of data from a given combination of experiments to identify parameters in a given flux depends on its ability to satisfy the conditions determined for practical identifiability of that parameter. For instance, identification of $V_1^{max}$ and $K_1^{ac}$ requires that the data from two different experiments used for parameter estimation must have distinct acetate concentrations as well as different uptake fluxes ($v_{1,j}$) between them (Equation \ref{eq:v1_par}). Thus, the informativeness of the experiments (changes in the input acetate concentration and $v_1$) for identifying parameters of $v_1$ is determined by the ability of the input change to effect a change in the measured value of $v_1$.	Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we say that identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for $v_1$.

	Although we were able to determine the requirements for identifying $v_1$ and $v_2$ based on the closed form expressions of the parameters, this may not be possible when a complex nonlinear rate law is used to model a flux. Using the Convenience Kinetic rate law to model $v_3$ is a case in point. So, in order to determine the type of experiments that help satisfy identifiability conditions for parameters of any flux modeled by any usable rate law, we look at the distribution of the different types of experiments in all data combinations that can identify the parameters within a flux. Recall that we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Section \ref{sec:experiments} and Table \ref{tab:pval}).

	The contribution from different types of experiments towards identification of parameters in different fluxes is given in Figure \ref{fig:exp_info}. In Figure \ref{fig:exp_info}a and b, the contribution from different experiment types for identifying $V_1^{max}$ and $K_1^{ac}(ne)$ are given, respectively. The contribution of different experiments towards identifiability for $V_2^{max}$ (Figure \ref{fig:exp_info}c) and $K_2^{pep}$ (Figure \ref{fig:exp_info}d) are also shown.

@@

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1/v1_v2_v3_v5_ident}}
		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$ when there is no noise in the input experimental data. The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter. The total number of experiments used to identify parameters in $v_1$, $v_2$ and $v_5$ is two. The total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is 240, 421 different combinations are used to test identifiability for $v_5$, and 4896 combinations are used to test identifiability of parameters for $v_3$. Note that $v_3$ is modeled with three parameters and requires data from at least three different experiments, and also has two possible roots for each of its parameters. Results for only one of the two roots indicated by (1) is shown here.}\label{fig:ident}
	\end{figure}
